---
title: "Problem_Set_1"
author: "Kristopher C. Toll"
date: "January 26, 2018"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## R Programing exercises

1. Calculate the square root of 729

```{r}
b = sqrt(729)
print(b)
```

2. Create a new variable a with value 1947.0

```{r}
a = 1947.0
print(a)
```

3. Create a vector b containg number form 1 to 6 and find out it's class.

b is a numeric variable.

```{r}
b = seq(1, 6, by = 1)
b.1 <- class(b)
print(c(b, b.1))
```

4. Create a vector c containg following mixed elements

```{r}
c = c(1, "a", 2, "b")
print(c)

```

  (a) Find out its class
  It is a character variable

```{r}
class(c)
```

  (b) Get the length of the vector.
  The length is four
  
```{r}
length(c) # Figuring out the length of c
```

  (c) Get the 2nd and 3rd elements, which is "a" and "2".
  
```{r}
print(c[2]) # Printing the 2nd element
print(c[3]) # Printing the 3rd element
```

5. Create a vector d containing following elements c(1, 2, NA, 4, 5, 6, NA, NA, NA, 10)
Remove missing valuse from d

```{r}
d = c(1, 2, NA, 4, 5, 6, NA, NA, NA, 10)
d = as.numeric(na.omit(d)) # Removing NA valuse and converting the vector to a numeric one
print(d)
```

6. Create a vector of values of e^x cos 3 at x = 3, 3.1, 3.2, ... 6

```{r}
x = seq(3, 6, by = 0.1)
x.1 = exp(x) * cos(x) # cosine is reading in x as a radian
print(x.1)
#x.2 = exp(x) * cos(x/180) # this funtions will read it in as degrees
#print(x.2)
```

7. Calculate $$\sum_{i = 10}^{100} (i^3 + 4*i^2) $$

```{r}
s = seq(10, 100, by = 1) # Creating the sequence
s.1 <- s^3 + 4 * s^2 # transforming the sequence
s.2 = sum(s.1) # Calculating the sum of the transformed sequence
print(s.2)
```

8. Execute the folling line wich create two vectors of random integers that are chosen with replacement from the integers 0, 1, ...999. Both vectors have length 250.

```{r}
x <- sample(0:999, size = 250, replace = TRUE)
y <- sample(0:999, size = 250, replace = TRUE)
```

  (a) Pick out the values in Y which are > 600
  
```{r}
y.1 <- as.numeric(y>600)
print(y>600)
```