---
title: "Problem_Set_1"
author: "Kristopher C. Toll"
date: "January 26, 2018"
output:
  word_document: default
  pdf_document: default
  html_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# R Programing exercises

###1. Calculate the square root of 729

```{r}
b = sqrt(729)
print(b)
```

###2. Create a new variable a with value 1947.0

```{r}
a = as.integer(1947.0)
print(a)
```

###3. Create a vector b contaning number form 1 to 6 and find out it's class.

b is a numeric variable.

```{r}
b = seq(1, 6, by = 1)
b.1 <- class(b)
print(c(b, b.1))
```

###4. Create a vector c containg following mixed elements

```{r}
c = c(1, "a", 2, "b")
print(c)

```

  (a) Find out its class.
  It is a character variable

```{r}
class(c)
```

  (b) Get the length of the vector.
  The length is four
  
```{r}
length(c) # Figuring out the length of c
```

  (c) Get the 2nd and 3rd elements, which is "a" and "2".
  
```{r}
print(c[2]) # Printing the 2nd element
print(c[3]) # Printing the 3rd element
```

###5. Create a vector d containing following elements c(1, 2, NA, 4, 5, 6, NA, NA, NA, 10)
Remove missing values from d

```{r}
d = c(1, 2, NA, 4, 5, 6, NA, NA, NA, 10)
d = as.numeric(na.omit(d)) # Removing NA valuse and converting the vector to a numeric one
print(d)
```

###6. Create a vector of values of $e^x$ cos 3 at x = 3, 3.1, 3.2, ... 6

```{r}
x = seq(3, 6, by = 0.1)
x.1 = exp(x) * cos(x) # cosine is reading in x as a radian unit
print(x.1)
#x.2 = exp(x) * cos(x/180) # this funtions will read it in as degrees
#print(x.2)
```

###7. Calculate $\sum_{i = 10}^{100} (i^3 + 4*i^2)$

```{r}
s = seq(10, 100, by = 1) # Creating the sequence
s.1 <- s^3 + 4 * s^2 # transforming the sequence
s.2 = sum(s.1) # Calculating the sum of the transformed sequence
print(s.2)
```

###8. Execute the following line which create two vectors of random integers that are chosen with replacement from the integers 0, 1, ...999. Both vectors have length 250.

```{r}
x <- sample(0:999, size = 250, replace = TRUE)
y <- sample(0:999, size = 250, replace = TRUE)
```

  (a) Pick out the values in Y which are > 600
  
```{r}
y.1 <- subset(y, y>600) # use the subset comand to pull out elements of a vector
print(y.1)
```

  (b) How many values in y are within 200 of the maximum value of the terms in y?
  
```{r}
y.2 <- subset(y, y >= max(y)-200)
print(length(y.2))
```


  (c) Create the vector e
  
```{r}
e <- abs(x-mean(x))^(1/2)
print(e)
```

  (d) Create another vector $(y_2-x_1, y_3-x_2, ... y_n - x_{n-1})$

```{r}
n=2:length(y)
d <- (y[n]-x[n-1]) # It worked!
print(d)
```

###9. In this exercise, we will consider a quadratic equations of the form $(y = \beta_0 + \beta_1 x + \beta_2 x^2)$. Create a vector of coefficients for a quadratic equations.

```{r}
coeffs <- sample(-20:20, size = 3, replace = TRUE)
```

(a) Determine the length of the object coeffs.

```{r}
print(length(coeffs))
```

(b) Create 200 values of x from a regularly spaced vector between -3 and 3

```{r}
x <- seq(from = -3, to = 3, length.out = 200)
```

(c) Now obtain the value of the quadratic function (y) at each value of x

```{r}
y = coeffs[1] + coeffs[2] * x + coeffs[3] * x^2
```

(d) Construct the plot

```{r}
plot(y, main = "Quadratic form of y")
```

###10. Without using R, determine the result of the following computation

$$ x <- c(1,2,3) $$

$$ x[1]/x[2]^3-1+2*x[3]-x[2-1] = 1/2^3-1+2*3-1 = 4.125 $$

###11. Create the following matrix with 15 rows

```{r}
A = matrix(c(rep(c(10, -5, 10), times = 15)), nrow = 15, byrow = TRUE)
print(A)

A.1 = A # Copy the matrix
A.1[,3] = A.1[,1] + A.1[,2] # rewrite the 3 column as a sum of the first two
print(A.1)
```

###12. Create a function that given two number will return he sum of those two number

```{r}
add <- function(a,b){
  c = a + b
  return(c)
}

add(5,10)
```

###13. Create a function that given a vector and an integer will return how many times the integer appears inside the vector

```{r}
count <- function(x, int){
  y <- vector()
  for(i in 1:length(x)){
    ifelse(x[i] == int, y[i] <- 1, y[i] <- 0)
  }
  z = sum(y)
  return(z)
}

# testing the function

x <- c(4,5, 6, 6, 7, 8) 
count(x = x, int = 6) # The argument should return a 2 given vector x
```

###14. Create a function that given an integer vector $(z_1, z_2, ... , z_n)$ will return $(z_1, z_1^2, ..., z_n^n)$

```{r}
zsquared <- function(x){
  z <- numeric(length(x))
  for(i in 1:length(x)){
    z[i] <- x[i]^i
  }
  return(z)
}

# Testing the function

x <- c(2, 2, 3)
zsquared(x) # should return a 2, 4, and 27 given vector x
```

###15. Create a piecewise function

```{r}
piecewise <- function(x){
  y <- numeric(length(x))
  for(i in 1:length(x)){
    if(x[i] < 0 ){
      y[i] = (x[i]^2 + 2 * x[i] + 3)
    }
    else if(x[i] >= 0 & x[i] < 2 ){
      y[i] = (x[i] + 3)
    }
    else if( 2 <= x[i]){
      y[i] = (x[i]^2 + 4 * x[i] - 7)
    }
  }
  return(y)
}

# Testing the piecewise function

x <-seq(-3, 3, length = 100)
y <- piecewise(x)

plot(x, y, main = "Piecewise function for #15", pch = "*", col = 617, bg = 456)
```

# Theory

### Problem 1: 
Show that for $Y_t$, $E(Y_t^2) = \sigma^2$ with $E(Y_t) = \mu_t = 0$ and $Var(Y_t) = \gamma_Y (0) = \sigma^2$



$$Var(Y_t) = \sigma^2 = E[(Y_t - \mu_t)^2] = E[Y_t^2 - 2 Y_t \mu + \mu_t^2]$$
$$\sigma^2 = E[Y_t^2] - 2 \mu_t E[Y_t] + E[\mu_t^2] = E[Y_t^2] - 2 \mu_t * 0 + 0 $$
$$\sigma^2 = E[Y_t^2] $$

### Problem 2:
Show that the autocovariance function can be written as $\gamma_y (s,t) = E(Y_s = \mu_s)(Y_t - \mu_t) = E(Y_s Y_t) - \mu_s \mu_t$ where $E(Y_t) = \mu_t$ and $E(Y_s) = \mu_s$

$$\gamma_y (s,t) = E(Y_s - \mu_s)(Y_t - \mu_t) = E(Y_s Y_t - Y_s \mu_t - Y_t \mu_s + \mu_t \mu_s)$$
$$\gamma_y (s,t) = E(Y_s Y_t) - E(Y_s \mu_t) - E(Y_t \mu_s) + E(\mu_t \mu_s) = E(Y_s Y_t) - \mu_t E(Y_s) - \mu_s E(Y_t) + \mu_t \mu_s $$
$$\gamma_y (s,t) =E(Y_s Y_t) - \mu_t \mu_s - \mu_s \mu_t + \mu_t \mu_s = E(Y_s Y_t) - 2\mu_t \mu_s + \mu_t \mu_s = E(Y_s Y_t) - \mu_t \mu_s$$
$$\gamma_y (s,t) = E(Y_s Y_t) - \mu_t \mu_s$$
